---
title: "ImportIch"
author: "Catarina Pien"
date: "1/4/2022"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Packages
```{r}
library(dplyr)
library(readxl)
library(lubridate)
```

#Ichthyoplankton

## New Template format

Read in file from Data_notprocessed folder.
```{r}
ich_files <- list.files("Data_Ich/Data_notprocessed", pattern = "*.xlsx", full.names = TRUE)
ich_comb <- lapply(ich_files, read_excel, sheet = "New_template_blank") %>%bind_rows()
```


Pivot lengths
```{r}
ich_long <- tidyr::pivot_longer(ich_comb, cols=LENGTH_1:LENGTH_30, values_to = "Length_mm", names_to = "Length_number")# %>%
 # filter(!is.na(Length_mm))
```

Re-organize
Add Larval columns
```{r}
ich_clean <- ich_long %>%
  janitor::clean_names() %>%
  mutate(measuring_program = "YBFMP",
         date,
         time2 = paste0(hour(time), ":", minute(time)),
         sam_code = sam_wdl_code,
         short_name = "ICH",
         attribute = "Ichthyoplankton", 
         total_length = "",
         life_stage = replace(life_stage, life_stage == "Preflexion", 
                              "L-Preflexion"),
         letter = substr(life_stage, 1,1),
         larval_life_stage = ifelse(letter == "L", life_stage, ""),
         life_stage_2 = ifelse(letter == "L", "L",
                             ifelse(letter == "Y", "YSL",
                                    ifelse(letter == "J", "J", NA)))) %>%
  select(measuring_program,
         date,
         time = time2,
         station,
         sam_code,
         short_name,
         attribute,
         observable = taxon_name,
         count,
         total_length,
         fork_length = length_mm,
         life_stage = life_stage_2,
         larval_life_stage)
str(ich_clean)
```

Write
```{r}
start <- format(first(ich_clean$date), "%Y%m%d")
end <- format(last(ich_clean$date), "%Y%m%d")
write.csv(ich_clean, paste0("Written_data/Ich_", start, "-", end, ".csv"), row.names = FALSE)
```

Once done, move the data out of the Data_not processed folder! Don't really need to save the file in here, as it should also be in the Biological Data folder, but can keep it until the next round in case there are formatting issues. 
































## For Old Data, not using new template
This is the first format, where time and date were "General" in Excel file.
```{r}
ich_files <- list.files("Data_Ich", pattern="*.xlsx", full.names=TRUE)

ich_df <- lapply(ich_files, read_excel, sheet = "Taxonomy Data", skip=6, col_types = c("text", "text", "skip", "guess", "skip", "skip", "text", "text", "text", "text", "text", "text", "text", "text","text", "text","text", "text","text", "text","text", "text","text", "text","text", "text","text", "text","text", "text","text", "text","text", "text","text", "text","text", "text","text", "text", "skip", "skip", "skip", "skip", "skip", "skip", "skip", "skip", "skip"))

ich_df1 <- lapply(ich_df, `[`, -c(1:1),)

ich_comb_1 <- bind_rows(ich_df1, .id = "dataset")%>%
  filter(!is.na(Taxon)) %>%
  mutate(Date = lubridate::mdy(Date),
         Time = strptime(Time, format="%H:%M")  %>%
           strftime(Time, format = "%H:%M")) 
str(ich_comb_1)
```

This is the second format, where time was a time and date was a date
```{r}
ich_files2 <- list.files("Data_Ich/Format2", pattern="*.xlsx", full.names=TRUE)

ich_df2 <- lapply(ich_files2, read_excel, sheet = "Taxonomy Data", skip=6, col_types = c("text", "text", "skip", "guess", "skip", "skip", "text", "text", "text", "text", "text", "text", "text", "text","text", "text","text", "text","text", "text","text", "text","text", "text","text", "text","text", "text","text", "text","text", "text","text", "text","text", "text","text", "text","text", "text", "skip", "skip", "skip", "skip", "skip", "skip", "skip", "skip", "skip"))

ich_df2_times <- ich_df2 


today<-as.POSIXct('2012-01-23 00:00:00 EST')
today + (3600*x)


ich_df_f2 <- lapply(ich_df2, `[`, -c(1:1),)
ich_comb_2 <- bind_rows(ich_df2, .id = "dataset") %>%
  filter(!is.na(Taxon))%>%
  mutate(Date = lubridate::ymd(Date),
         Time = as.numeric(Time),
         Hour = floor(Time * 24),
         Minute = (((Time*24) - Hour) * 60),
         Time2 = paste0(Hour, ":", Minute),
         Time2 = strptime(Time2, format="%H:%M")  %>%
           strftime(Time2, format = "%H:%M")) 

ich_comb_3 <- ich_comb_2 %>%select(-Time, -Hour, -Minute) %>%
  rename(Time = Time2) %>%
  select(dataset, Station, Date, Time, everything())

str(ich_comb_3)
```

Combine
```{r}
ich_comb <- rbind(ich_comb_1, ich_comb_3)
```

Pivot lengths
```{r}
ich_long <- tidyr::pivot_longer(ich_comb, cols=9:38, values_to = "Length_mm") %>%
  filter(!is.na(Length_mm))
```

Re-organize
Add Larval columns
```{r}
ich_clean <- ich_long %>%
  janitor::clean_names() %>%
  rename(life_stage = life_stage_y_l_j) %>%
  mutate(measuring_program = "YBFMP",
         date,
         sam_code = "",
         short_name = "ICH",
         attribute = "Ichthyoplankton", 
         total_length = "",
         life_stage = replace(life_stage, life_stage == "Preflexion", 
                              "L-Preflexion"),
         letter = substr(life_stage, 1,1),
         larval_life_stage = ifelse(letter == "L", life_stage, ""),
         life_stage_2 = ifelse(letter == "L", "L",
                             ifelse(letter == "Y", "YSL",
                                    ifelse(letter == "J", "J", NA)))) %>%
  select(measuring_program,
         date,
         time,
         station,
         sam_code,
         short_name,
         attribute,
         observable = taxon,
         count,
         total_length,
         fork_length = length_mm,
         life_stage = life_stage_2,
         larval_life_stage)
str(ich_clean)
```

Write
```{r}
start <- format(first(ich_clean$date), "%Y%m%d")
end <- format(last(ich_clean$date), "%Y%m%d")
write.csv(ich_clean, paste0("Written_data/Ich_", start, "-", end, ".csv"), row.names = FALSE)
```
