---
title: "ImportIch"
author: "Catarina Pien"
date: "1/4/2022"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Packages
```{r}
library(dplyr)
library(readxl)
library(lubridate)
library(janitor)
```

#Ichthyoplankton

## New Template format

Before you start, move old data out of the Data_not processed folder! Don't really need to save the file in the repo, as it should also be in the Biological Data folder, but can keep it for a bit in case it needs to be re-processed. 


Read in file from Data_notprocessed folder. Can change the folder if needed. Should be able to read multiple files but haven't tested with the new code.
```{r}
ich_files <- list.files("Data_Ich/Data_notprocessed", pattern = "*.xlsx", full.names = TRUE)
ich_comb <- lapply(ich_files, read_excel, sheet = "New_template_blank") %>%
  bind_rows() %>%
  mutate(TAXON_NAME = ifelse(TAXON_NAME == "N/A", "No Catch", TAXON_NAME))
```

Pivot lengths

Please check this once there are examples with catch. I ran this for an example with no catch. 
```{r}
ich_long <- tidyr::pivot_longer(ich_comb, cols=LENGTH_1:LENGTH_30, values_to = "Length_mm", names_to = "Length_number") %>%
  mutate(Length_mm = ifelse(Length_number == "LENGTH_1" & TAXON_NAME == "No Catch", "NOCATCH", Length_mm))%>%
 filter(!is.na(Length_mm))
```

Re-organize
Add Larval columns
```{r}
ich_clean <- ich_long %>%
  janitor::clean_names() %>%
  mutate(measuring_program = "YBFMP",
         date,
         time2 = paste0(hour(time), ":", minute(time)),
         sam_code = ifelse(is.na(sam_wdl_code), "", sam_wdl_code),
         short_name = "ICH",
        dwr_sample_id = ifelse(is.na(dwr_sample_id), paste0(format(date, "%Y%m%d"), "_", hour(time), minute(time), "_", station, "_", short_name), dwr_sample_id),
           attribute = "Ichthyoplankton", 
         total_length = "",
         life_stage = replace(life_stage, life_stage == "Preflexion", 
                              "L-Preflexion"),
         letter = substr(life_stage, 1,1),
         larval_life_stage = ifelse(letter == "L", life_stage, ""),
         life_stage_2 = ifelse(letter == "L", "L",
                             ifelse(letter == "Y", "YSL",
                                    ifelse(letter == "J", "J", NA)))) %>%
  select(measuring_program,
         date,
         time = time2,
         station,
         sam_code,
         sample_id = dwr_sample_id,
         short_name,
         attribute,
         observable = taxon_name,
         count,
         total_length,
         fork_length = length_mm,
         life_stage = life_stage_2,
         larval_life_stage) %>%
  mutate(fork_length = ifelse(fork_length == "NOCATCH", NA, fork_length))
str(ich_clean)
```

Write file
```{r}
start <- format(first(ich_clean$date), "%Y%m%d")
end <- format(last(ich_clean$date), "%Y%m%d")
write.csv(ich_clean, paste0("Written_data/Ich_", start, "-", end, ".csv"), row.names = FALSE)
```



































## For Old Data, not using new template
This is the first format, where time and date were "General" in Excel file.
```{r}
ich_files <- list.files("Data_Ich", pattern="*.xlsx", full.names=TRUE)

ich_df <- lapply(ich_files, read_excel, sheet = "Taxonomy Data", skip=6, col_types = c("text", "text", "skip", "guess", "skip", "skip", "text", "text", "text", "text", "text", "text", "text", "text","text", "text","text", "text","text", "text","text", "text","text", "text","text", "text","text", "text","text", "text","text", "text","text", "text","text", "text","text", "text","text", "text", "skip", "skip", "skip", "skip", "skip", "skip", "skip", "skip", "skip"))

ich_df1 <- lapply(ich_df, `[`, -c(1:1),)

ich_comb_1 <- bind_rows(ich_df1, .id = "dataset")%>%
  filter(!is.na(Taxon)) %>%
  mutate(Date = lubridate::mdy(Date),
         Time = strptime(Time, format="%H:%M")  %>%
           strftime(Time, format = "%H:%M")) 
str(ich_comb_1)
```

This is the second format, where time was a time and date was a date
```{r}
ich_files2 <- list.files("Data_Ich/Format2", pattern="*.xlsx", full.names=TRUE)

ich_df2 <- lapply(ich_files2, read_excel, sheet = "Taxonomy Data", skip=6, col_types = c("text", "text", "skip", "guess", "skip", "skip", "text", "text", "text", "text", "text", "text", "text", "text","text", "text","text", "text","text", "text","text", "text","text", "text","text", "text","text", "text","text", "text","text", "text","text", "text","text", "text","text", "text","text", "text", "skip", "skip", "skip", "skip", "skip", "skip", "skip", "skip", "skip"))

ich_df2_times <- ich_df2 


today<-as.POSIXct('2012-01-23 00:00:00 EST')
today + (3600*x)


ich_df_f2 <- lapply(ich_df2, `[`, -c(1:1),)
ich_comb_2 <- bind_rows(ich_df2, .id = "dataset") %>%
  filter(!is.na(Taxon))%>%
  mutate(Date = lubridate::ymd(Date),
         Time = as.numeric(Time),
         Hour = floor(Time * 24),
         Minute = (((Time*24) - Hour) * 60),
         Time2 = paste0(Hour, ":", Minute),
         Time2 = strptime(Time2, format="%H:%M")  %>%
           strftime(Time2, format = "%H:%M")) 

ich_comb_3 <- ich_comb_2 %>%select(-Time, -Hour, -Minute) %>%
  rename(Time = Time2) %>%
  select(dataset, Station, Date, Time, everything())

str(ich_comb_3)
```

Combine
```{r}
ich_comb <- rbind(ich_comb_1, ich_comb_3)
```

Pivot lengths
```{r}
ich_long <- tidyr::pivot_longer(ich_comb, cols=9:38, values_to = "Length_mm") %>%
  filter(!is.na(Length_mm))
```

Re-organize
Add Larval columns
```{r}
ich_clean <- ich_long %>%
  janitor::clean_names() %>%
  rename(life_stage = life_stage_y_l_j) %>%
  mutate(measuring_program = "YBFMP",
         date,
         sam_code = "",
         short_name = "ICH",
         attribute = "Ichthyoplankton", 
         total_length = "",
         life_stage = replace(life_stage, life_stage == "Preflexion", 
                              "L-Preflexion"),
         letter = substr(life_stage, 1,1),
         larval_life_stage = ifelse(letter == "L", life_stage, ""),
         life_stage_2 = ifelse(letter == "L", "L",
                             ifelse(letter == "Y", "YSL",
                                    ifelse(letter == "J", "J", NA)))) %>%
  select(measuring_program,
         date,
         time,
         station,
         sam_code,
         short_name,
         attribute,
         observable = taxon,
         count,
         total_length,
         fork_length = length_mm,
         life_stage = life_stage_2,
         larval_life_stage)
str(ich_clean)
```

Write
```{r}
start <- format(first(ich_clean$date), "%Y%m%d")
end <- format(last(ich_clean$date), "%Y%m%d")
write.csv(ich_clean, paste0("Written_data/Ich_", start, "-", end, ".csv"), row.names = FALSE)
```
