---
title: "ImportDriftData"
author: "Catarina Pien"
date: "12/1/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Packages
```{r}
library(dplyr)
library(readxl)
library(lubridate)
library(janitor)
```

# Drift data

## Current Code
Before you start, move old data out of the Data_not processed folder! Don't really need to save the file in the repo, as it should also be in the Biological Data folder, but can keep it for a bit in case it needs to be re-processed. 

* List the folder where all the data are found. If there is new data, can always make a new folder. 
```{r}

drift_files <- list.files("Data_Drift/Data_notprocessed", pattern="*.xlsx", full.names=TRUE)

# Need to check original data files to make sure the first 8 rows are to be skipped. There was one file that was 9 I had to modify.
ldf <- lapply(drift_files, read_excel, sheet = "CADWR template") %>%bind_rows()

```

Clean columns
Have to check if 
```{r}
glimpse(drift_data_clean)
# want date to be in dttm and time to be in chr. otherwise play with some of those commented out datetime code lines.

drift_clean <- ldf %>%
  janitor::clean_names() %>%
  mutate(measuring_program = "YBFMP",
         dwr_sample_id = ifelse(is.na(dwr_sample_id), paste0(format(date, "%Y%m%d"), "_", hour(hm(time)), minute(hm(time)), "_", station, "_", short_name), dwr_sample_id),
        # time2 = paste0(hour(time), ":", minute(time)),
        # date = as.Date(as.numeric(as.character(date)), origin = "1899-12-30"),
        # Time = as.numeric(time),
        # Hour = floor(Time * 24),
        # Minute = (((Time*24) - Hour) * 60),
        # Time2 = paste0(Hour, ":", Minute),
        # time = strptime(Time2, format="%H:%M")  %>%
        #   strftime(Time2, format = "%H:%M"), 
         date = date_col,
         sam_code = ifelse(is.na(sam_wdl_code), "", sam_wdl_code),
         short_name = "DRIFT",
         attribute = "Drift Invertebrates") %>%
  select(measuring_program,
         date,
         time,
         station,
         sam_code,
         sample_id= dwr_sample_id,
         short_name,
         attribute,
         observable = taxon_name,
         category = origin,
         count = abundance,
         life_stage,
         wet_mass,
         dry_mass,
         lab_comments = lab_com,
         condition)
str(drift_clean)
  
```

Write data
```{r}
first <- format(first(drift_clean$date), "%Y%m%d")
last <- format(last(drift_clean$date), "%Y%m%d")
write.csv(drift_clean, paste0("Written_data/Drift_", first, "-", last, ".csv"), row.names = FALSE)
```

















## Old Code for historical data, multiple files

### Read in files, get the column names lined up, bind to data frame

* List the folder where all the data are found. If there is new data, can always make a new folder. 
* First make sure that all data files are formatted in the same way. 8 blank rows at the top, column name, then two descriptive rows. Can modify the "skip" for blank rows, and the "1:2" based on any differences. 
```{r}

drift_files <- list.files("Data_Drift", pattern="*.xlsx", full.names=TRUE)

# Need to check original data files to make sure the first 8 rows are to be skipped. There was one file that was 9 I had to modify.
ldf <- lapply(drift_files, read_excel, sheet = "Taxonomy Data", skip=8, col_types = c("text", "text", "text", "text", "skip", "skip", "skip", "skip", "skip", "text", "text", "text", "text", "text", "text", "text", "skip", "skip", "text", "text", "text", "text", "skip", "skip", "text", "skip", "skip", "skip", "skip", "skip", "skip","skip", "skip", "skip","skip", "skip", "skip", "skip", "skip", "skip","skip", "skip", "skip","skip", "skip", "skip" ))
ldf1 <- lapply(ldf, `[`, -c(1:2),) #remove two extra rows that are descriptive rows, but still want to keep column titles.

#
```

New template
```{r}
drift_files <- list.files("Data_Drift", pattern="*.xlsx", full.names=TRUE)

ldf1 <- lapply(drift_files, read_excel, sheet = "CADWR template", 
              range = cell_cols(1:25),
  col_types = c("text", "text", "text", "text", "text","text", 
                "skip", "text", "skip", "skip", "skip", "text", 
                "text", "text", "numeric", "text", "text", "skip", 
                "skip", "skip", "numeric", "text", "text", "text", "text"))
```



## Clean up names in each file so that they line up with each other. There were some inconsistencies in the original.
```{r}
f_name_change <- function(x) {
  x <- janitor::clean_names(x)
  #names(x)[2:6] <- c("station", "time", "date", "taxon", "life_stage")
  # other way to convert date
  # x <- x %>%
  #  mutate(date = janitor::excel_numeric_to_date(as.numeric(as.character(date)), date_system = "modern")) 
  return(x)
}

# apply function
ldf2 <- lapply(ldf1, f_name_change)
```

## Combine into a data frame
```{r}
# combine into data frame
drift_comb <- bind_rows(ldf2, .id = "dataset")
str(drift_comb)
```

## Select columns, rename
```{r}
drift_clean <- drift_comb %>%
  mutate(measuring_program = ifelse(study == "YB", "YBFMP", "NDFA"),
         date = as.Date(as.numeric(as.character(date)), origin = "1899-12-30"),
         Time = as.numeric(time),
         Hour = floor(Time * 24),
         Minute = (((Time*24) - Hour) * 60),
         Time2 = paste0(Hour, ":", Minute),
         time = strptime(Time2, format="%H:%M")  %>%
           strftime(Time2, format = "%H:%M"), 
         sam_code = "",
         short_name = "DRIFT",
         attribute = "Drift Invertebrates") %>%
  select(measuring_program,
         date,
         time,
         station,
         sam_code,
         short_name,
         attribute,
         observable = taxon_name,
         category = origin,
         count = abundance,
         life_stage,
         wet_mass,
         dry_mass,
         lab_comments = lab_com,
         condition)
str(drift_clean)
```

Write
```{r}
first <- format(first(drift_clean$date), "%Y%m%d")
last <- format(last(drift_clean$date), "%Y%m%d")
write.csv(drift_clean, paste0("Written_data/Drift_", first, "-", last, ".csv"), row.names = FALSE)
```


